---
title: "Estimating the intensive and extensive margin of trade"
output: rmarkdown::html_vignette
bibliography: lit.bib
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Estimating the intensive and extensive margin of trade}
  %\VignetteEncoding{UTF-8}
---


## Introduction

In econometrics, fixed effects models are popular to control for unobserved heterogeneity in data sets with a panel structure. In non-linear models this is usually done by including a dummy variable for each level of a fixed effects category. This approach can quickly become infeasible if the number of levels and/or fixed effects categories increases (either due to memory or time limitations).

Two examples of this limitation can be found in the literature on international trade. Firstly, when it comes to estimating the determinants of trade volumes (intensive margin) or secondly, when it comes to the export decisions (extensive margin). In order to demonstrate the functionality and capablilities of `alpaca`, we replicate parts of @lwyz19 as an example for the intensive margin and extend it to the extensive margin. Our example is based on the panel data set of 2,973,168 bilateral trade flows used by @gr16. For our estimations we specify a three-way error component with 56,608 levels.

## Data preparation

The data set is available either from [Andrew Rose's website][1] or from [sciencedirect][2]. We use the same variable names as @gr16 such that we are able to compare our summary statistics with the ones provided in their *Stata* log-files.

[1]: http://faculty.haas.berkeley.edu/arose/RecRes.htm
[2]: https://www.sciencedirect.com/science/article/pii/S0014292116300630#ec0005

```{r, eval = FALSE}
# Import the data set
library(haven)
library(data.table)
cudata <- read_dta("dataaxj1.dta")
setDT(cudata)

# Subsetting relevant variables
var.nms <- c("exp1to2", "custrict11", "ldist", "comlang", "border", "regional",
             "comcol", "curcol", "colony", "comctry", "cuwoemu", "emu", "cuc",
             "cty1", "cty2", "year", "pairid")
cudata <- cudata[, ..var.nms]

# Generate identifiers required for structural gravity
cudata[, pairid := factor(pairid)]
cudata[, exp.time := interaction(cty1, year)]
cudata[, imp.time := interaction(cty2, year)]

# Generate dummies for disaggregated currency unions
cudata[, cuau := as.integer(cuc == "au")]
cudata[, cube := as.integer(cuc == "be")]
cudata[, cuca := as.integer(cuc == "ca")]
cudata[, cucf := as.integer(cuc == "cf")]
cudata[, cucp := as.integer(cuc == "cp")]
cudata[, cudk := as.integer(cuc == "dk")]
cudata[, cuea := as.integer(cuc == "ea")]
cudata[, cuec := as.integer(cuc == "ec")]
cudata[, cuem := as.integer(cuc == "em")]
cudata[, cufr := as.integer(cuc == "fr")]
cudata[, cugb := as.integer(cuc == "gb")]
cudata[, cuin := as.integer(cuc == "in")]
cudata[, cuma := as.integer(cuc == "ma")]
cudata[, cuml := as.integer(cuc == "ml")]
cudata[, cunc := as.integer(cuc == "nc")]
cudata[, cunz := as.integer(cuc == "nz")]
cudata[, cupk := as.integer(cuc == "pk")]
cudata[, cupt := as.integer(cuc == "pt")]
cudata[, cusa := as.integer(cuc == "sa")]
cudata[, cusp := as.integer(cuc == "sp")]
cudata[, cuua := as.integer(cuc == "ua")]
cudata[, cuus := as.integer(cuc == "us")]
cudata[, cuwa := as.integer(cuc == "wa")]
cudata[, cuwoo := custrict11]
cudata[cuc %in% c("em", "au", "cf", "ec", "fr", "gb", "in", "us"), cuwoo := 0L]

# Set missing trade flows to zero
cudata[is.na(exp1to2), exp1to2 := 0.0]

# Construct binary and lagged dependent variable for the extensive margin
cudata[, trade := as.integer(exp1to2 > 0.0)]
cudata[, ltrade := shift(trade), by = pairid]
```


## Estimating the intensive margin of trade

After preparing the data, we show how to replicate column 3 of table 2 in @lwyz19. In addition to coefficients and robust standard errors, the authors also report standard errors clustered by exporter, importer, and time.

If we want `feglm()` to report standard errors that are clustered by variables, which are not already part of the model itself, we have to additionally provide them using the third part of the `formula` interface. In this example, we have to additionally add identifiers for exporters (`cty1`), importers (`cty2`), and time (`year`). 

First we report robust standard errors indicated by the option `"sandwich"` in `summary()`.
```{r, eval = FALSE}
mod <- feglm(exp1to2 ~ emu + cuwoo + cuau + cucf + cuec + cufr + cugb + cuin + cuus +
               regional + curcol | exp.time + imp.time + pairid | cty1 + cty2 + year, cudata,
             family = poisson())
summary(mod, "sandwich")
```

```{r, eval = FALSE}
## poisson - log link
## 
## exp1to2 ~ emu + cuwoo + cuau + cucf + cuec + cufr + cugb + cuin + 
##     cuus + regional + curcol | exp.time + imp.time + pairid | 
##     cty1 + cty2 + year
## 
## Estimates:
##           Estimate Std. error z value Pr(> |z|)    
## emu       0.048895   0.010277   4.758  1.96e-06 ***
## cuwoo     0.765988   0.053272  14.379   < 2e-16 ***
## cuau      0.384469   0.118832   3.235   0.00121 ** 
## cucf     -0.125608   0.099674  -1.260   0.20760    
## cuec     -0.877318   0.083451 -10.513   < 2e-16 ***
## cufr      2.095726   0.062952  33.291   < 2e-16 ***
## cugb      1.059957   0.034680  30.564   < 2e-16 ***
## cuin      0.169745   0.147029   1.154   0.24830    
## cuus      0.018323   0.021530   0.851   0.39473    
## regional  0.159181   0.008714  18.267   < 2e-16 ***
## curcol    0.386882   0.046827   8.262   < 2e-16 ***
## ---
## Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
## 
## residual deviance= 35830778.62,
## null deviance= 2245707302.17,
## n= 1610165, l= [11227, 11277, 34104]
## 
## ( 1363003 observation(s) deleted due to perfect classification )
## 
## Number of Fisher Scoring Iterations: 13 
```

We observe that roughly 1,4 million observations do not contribute to the identification of the structural parameters and we end up with roughly 1,6 million observations and 57,000 fixed effects.

Replicating the clustered standard errors is straightforward. We simply have to change the option to `"clustered"` and provide `summary` with the requested cluster dimensions.
```{r, eval = FALSE}
summary(mod, "clustered", cluster = ~ cty1 + cty2 + year)
```

```{r, eval = FALSE}
## poisson - log link
## 
## exp1to2 ~ emu + cuwoo + cuau + cucf + cuec + cufr + cugb + cuin + 
##     cuus + regional + curcol | exp.time + imp.time + pairid | 
##     cty1 + cty2 + year
## 
## Estimates:
##          Estimate Std. error z value Pr(> |z|)    
## emu       0.04890    0.09455   0.517   0.60507    
## cuwoo     0.76599    0.24933   3.072   0.00213 ** 
## cuau      0.38447    0.22355   1.720   0.08546 .  
## cucf     -0.12561    0.35221  -0.357   0.72137    
## cuec     -0.87732    0.29493  -2.975   0.00293 ** 
## cufr      2.09573    0.30625   6.843  7.75e-12 ***
## cugb      1.05996    0.23766   4.460  8.19e-06 ***
## cuin      0.16974    0.30090   0.564   0.57267    
## cuus      0.01832    0.05092   0.360   0.71898    
## regional  0.15918    0.07588   2.098   0.03593 *  
## curcol    0.38688    0.15509   2.495   0.01261 *  
## ---
## Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
## 
## residual deviance= 35830779,
## null deviance= 2245707302,
## n= 1610165, l= [11227, 11277, 34104]
## 
## ( 1363003 observation(s) deleted due to perfect classification )
## 
## Number of Fisher Scoring Iterations: 13
```


## Testing linear restrictions

Our package is also compatible with `linearHypothesis()` of the `car` package. In the next example we show how to test if all currency union effects except being in the EMU are jointly different from zero using a Wald test based on a clustered covariance matrix.
```{r, eval = FALSE}
library(car)
cus <- c("cuwoo", "cuau", "cucf", "cuec", "cufr", "cugb", "cuin", "cuus")
linearHypothesis(mod, cus, vcov. = vcov(mod, "clustered", cluster = ~ cty1 + cty2 + year))
```


```{r, eval = FALSE}
## Linear hypothesis test
## 
## Hypothesis:
## cuwoo = 0
## cuau = 0
## cucf = 0
## cuec = 0
## cufr = 0
## cugb = 0
## cuin = 0
## cuus = 0
## 
## Model 1: restricted model
## Model 2: exp1to2 ~ emu + cuwoo + cuau + cucf + cuec + cufr + cugb + cuin + 
##     cuus + regional + curcol | exp.time + imp.time + pairid | 
##     cty1 + cty2 + year
## 
## Note: Coefficient covariance matrix supplied.
## 
##   Df  Chisq Pr(>Chisq)    
## 1                         
## 2  8 96.771  < 2.2e-16 ***
## ---
## Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
```

## Estimating the extensive margin of trade

Now we turn to the estimation of the extensive margin. First we estimate a static probit model.
```{r, eval = FALSE}
mods <- feglm(trade ~ emu + cuwoo + cuau + cucf + cuec + cufr + cugb + cuin + cuus +
                regional + curcol | exp.time + imp.time + pairid, cudata,
              family = binomial("probit"))
summary(mods)
```

```{r, eval = FALSE}
## binomial - probit link
## 
## trade ~ emu + cuwoo + cuau + cucf + cuec + cufr + cugb + cuin + 
##     cuus + regional + curcol | exp.time + imp.time + pairid
## 
## Estimates:
##          Estimate Std. error z value Pr(> |z|)    
## emu       0.28257    0.12830   2.202  0.027633 *  
## cuwoo     0.56822    0.07686   7.393  1.44e-13 ***
## cuau      1.10887    0.28988   3.825  0.000131 ***
## cucf      0.13957    0.05263   2.652  0.008001 ** 
## cuec      0.05410    0.61908   0.087  0.930358    
## cufr      2.30667    0.24342   9.476   < 2e-16 ***
## cugb      0.18393    0.04272   4.305  1.67e-05 ***
## cuin      0.25151    0.15735   1.598  0.109947    
## cuus     -0.36144    0.07948  -4.547  5.43e-06 ***
## regional  0.06285    0.01787   3.518  0.000435 ***
## curcol   -0.89259    0.23332  -3.826  0.000130 ***
## ---
## Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
## 
## residual deviance= 681267.38,
## null deviance= 1947737.97,
## n= 1406505, l= [11227, 11273, 29630]
## 
## ( 1566663 observation(s) deleted due to perfect classification )
## 
## Number of Fisher Scoring Iterations: 7
```

To mitigate the incidental parameters problem (see @ns48) we apply the bias correction suggested by @hsw19. Note that the argument `panel.structure = "network"` is necessary to apply the appropriate bias correction.
```{r, eval = FALSE}
modsbc <- biasCorr(mods, panel.structure = "network")
summary(modsbc)
```

```{r, eval = FALSE}
## binomial - probit link
## 
## trade ~ emu + cuwoo + cuau + cucf + cuec + cufr + cugb + cuin + 
##     cuus + regional + curcol | exp.time + imp.time + pairid
## 
## Estimates:
##          Estimate Std. error z value Pr(> |z|)    
## emu       0.26909    0.12830   2.097  0.035957 *  
## cuwoo     0.50380    0.07686   6.555  5.58e-11 ***
## cuau      0.99443    0.28988   3.430  0.000603 ***
## cucf      0.12965    0.05263   2.463  0.013763 *  
## cuec      0.01616    0.61908   0.026  0.979177    
## cufr      2.08515    0.24342   8.566   < 2e-16 ***
## cugb      0.16730    0.04272   3.916  9.00e-05 ***
## cuin      0.21050    0.15735   1.338  0.180960    
## cuus     -0.33648    0.07948  -4.233  2.30e-05 ***
## regional  0.06350    0.01787   3.554  0.000379 ***
## curcol   -0.72566    0.23332  -3.110  0.001870 ** 
## ---
## Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
## 
## residual deviance= 681267.38,
## null deviance= 1947737.97,
## n= 1406505, l= [11227, 11273, 29630]
## 
## ( 1566663 observation(s) deleted due to perfect classification )
## 
## Number of Fisher Scoring Iterations: 7 
```

Because coefficients itself are not very meaningful, econometricians are usually interested in so called partial effects (also known as marginal or ceteris paribus effects). A commonly used statistic is the average partial effect. `alpaca` offers a post-estimation routine to estimate average partial effects and their corresponding standard errors. In the following the bias-corrected average partial effects are computed.
```{r, eval = FALSE}
apesbc <- getAPEs(modsbc, n.pop = nrow(cudata))
summary(apesbc)
```

```{r, eval = FALSE}
## Estimates:
##           Estimate Std. error z value Pr(> |z|)    
## emu       0.020386   0.018029   1.131  0.258178    
## cuwoo     0.038504   0.012516   3.076  0.002095 ** 
## cuau      0.076624   0.031081   2.465  0.013688 *  
## cucf      0.009758   0.008048   1.212  0.225349    
## cuec      0.001209   0.080143   0.015  0.987966    
## cufr      0.179129   0.050085   3.576  0.000348 ***
## cugb      0.012612   0.006424   1.963  0.049600 *  
## cuin      0.015905   0.023544   0.676  0.499333    
## cuus     -0.024633   0.011286  -2.183  0.029067 *  
## regional  0.004765   0.002633   1.810  0.070334 .  
## curcol   -0.051613   0.037436  -1.379  0.167982    
## ---
## Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
```

@hsw19 propose to estimate the extensive margin using a dynamic specification, where the lagged dependent variable is added to the explanatory variables. Again, a bias correction is necessary. Contrary to the one for static models, we need to additionally provide a bandwidth parameter (`L`) that is required for the estimation of spectral densities (see @hk11). @fw16 suggest to do a sensitivity analysis and try different values for `L` but not larger than four. Note that in this case the order of factors to be concentrated out, specified in the second part of the formula, is important (importer-/exporter-time identifiers first and pair identifier last).
```{r, eval = FALSE}
modd <- feglm(trade ~ ltrade + emu + cuwoo + cuau + cucf + cuec + cufr + cugb + cuin + cuus +
                regional + curcol | exp.time + imp.time + pairid, cudata,
              family = binomial("probit"))
moddbc <- biasCorr(modd, L = 2L, panel.structure = "network")
summary(moddbc)
```

```{r, eval = FALSE}
## binomial - probit link
## 
## trade ~ ltrade + emu + cuwoo + cuau + cucf + cuec + cufr + cugb + 
##     cuin + cuus + regional + curcol | exp.time + imp.time + pairid
## 
## Estimates:
##          Estimate Std. error z value Pr(> |z|)    
## ltrade    1.24341    0.00443 280.702   < 2e-16 ***
## emu       0.18274    0.13543   1.349   0.17724    
## cuwoo     0.41965    0.08064   5.204  1.95e-07 ***
## cuau      0.79096    0.30410   2.601   0.00930 ** 
## cucf      0.10147    0.05446   1.863   0.06244 .  
## cuec      0.17175    0.68499   0.251   0.80202    
## cufr      1.46319    0.24244   6.035  1.59e-09 ***
## cugb      0.10097    0.04451   2.268   0.02331 *  
## cuin      0.17055    0.16394   1.040   0.29819    
## cuus     -0.25668    0.08215  -3.124   0.00178 ** 
## regional  0.04656    0.01867   2.494   0.01262 *  
## curcol   -0.44021    0.25167  -1.749   0.08026 .  
## ---
## Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
## 
## residual deviance= 603826.72,
## null deviance= 1930471.73,
## n= 1394201, l= [11133, 11178, 29538]
## 
## ( 45048 observation(s) deleted due to missingness )
## ( 1533919 observation(s) deleted due to perfect classification )
## 
## Number of Fisher Scoring Iterations: 13 
```

Again we compute bias-corrected average partial effects to get interpretable quantities.
```{r, eval = FALSE}
apedbc <- getAPEs(moddbc, n.pop = nrow(cudata[!is.na(ltrade)]))
summary(apedbc)
```

```{r, eval = FALSE}
## Estimates:
##           Estimate Std. error z value Pr(> |z|)    
## ltrade    0.128362   0.000981 130.849   < 2e-16 ***
## emu       0.012547   0.016227   0.773   0.43942    
## cuwoo     0.029436   0.010014   2.940   0.00329 ** 
## cuau      0.057520   0.031869   1.805   0.07109 .  
## cucf      0.006920   0.006713   1.031   0.30261    
## cuec      0.011782   0.077696   0.152   0.87947    
## cufr      0.112661   0.051794   2.175   0.02962 *  
## cugb      0.006885   0.005096   1.351   0.17666    
## cuin      0.011697   0.019782   0.591   0.55430    
## cuus     -0.017074   0.009183  -1.859   0.06298 .  
## regional  0.003162   0.002185   1.447   0.14793    
## curcol   -0.029015   0.032597  -0.890   0.37342    
## ---
## Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
```

## References